On constructing permutation-fair dice

### 2018-04-15 postdoctoral/go-first.tex
more go-first progress

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newcommand{\rest}{\upharpoonright}
\newcommand{\term}{\textbf}
\begin{document}

@@

\begin{definition}
    Given a formula \(F:n\to S\) and \(m\leq n\), let \(F\rest m:m\to S\) be the 
    \term{restriction formula} where \((F\rest m)(k)=F(k)\) for \(k<m\).
\end{definition}

@@

\begin{definition}
    A \term{sample} for a formula \(F:n\to S\) is a function \(f:S\to n\) such
    that \(F(f(A))=A\) for all \(A\in S\).
\end{definition}

\begin{definition}
    The \term{winner} of a given sample is \(A\in S\) such that
    \(f(A)=\max\{f(B):B\in S\}\).
\end{definition}

\begin{definition}
    A formula \(F:n\to S\) is called \term{go-first-fair} if each \(A\in S\)
    is the winner of an equal number of samples of \(F\).
\end{definition}

\begin{definition}etc.\end{definition}

\begin{proposition}
    Suppose \(F:n\to S\) is a permutation-fair (resp. place-fair, go-first-fair)
    formula. Then \(F_T\) is permutation-fair (resp. place-fair, go-first-fair)
    for all \(T\subseteq S\).
\end{proposition}

\begin{theorem}
    Suppose \(F:n\to S\cup\{X\}\) is a go-first-fair formula such that 
    for each \(m\leq n\) where \(f(m)=X\), 
    \((F\rest m)_S\) is permutation-fair.
    Then \(F\) is permutation-fair.
\end{theorem}

@@

>>>>

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newcommand{\rest}{\upharpoonright}
\newcommand{\term}{\textbf}

\title{On constructing permutation-fair dice}
\author{Steven Clontz}
\begin{document}
\maketitle

@@

\begin{definition}
    Given a sequence \(F:n\to S\) and \(m\leq n\), let \(F\rest m:m\to S\) be the 
    \term{restriction sequence} where \((F\rest m)(k)=F(k)\) for \(k<m\).
\end{definition}

@@

\begin{definition}
    A \term{sample} for a formula \(F:n\to S\) is a function \(s:S\to n\) such
    that \(F(s(A))=A\) for all \(A\in S\). Let \(\sigma (F)\) collect
    all samples for the formula \(F\).
\end{definition}

\begin{definition}
    The \term{result} of a sample \(s:S\to n\) is a permutation \(r_s:|S|\to S\)
    of \(S\) such that \(i\leq j\) if and only if \(f(r_s(i))\leq f(r_s(j))\).
\end{definition}

So for the formula \(ABCACBA\), a sample might be defined
by \(f(A)=3,f(B)=2,f(C)=4\), yielding the result \(BAC\).
Note that the highest-valued die in the sample \(s\) is given by \(r_s(|S|-1)\).

\begin{definition}
    A formula \(F:n\to S\) is called \term{permutation-fair} if
    for all permutations \(p,q\) of \(S\),
    \(|\{s\in\sigma (F):r_s=p\}|=|\{s\in\sigma (F):r_s=q\}|\). 
\end{definition}

\begin{definition}
    A formula \(F:n\to S\) is called \term{go-first-fair} if
    for all \(A,B\in S\),
    \(|\{s\in\sigma (F):r_s(|S|-1)=A\}|=|\{s\in\sigma (F):r_s(|S|-1)=B\}|\). 
\end{definition}

\begin{proposition}
    Suppose \(F:n\to S\) is a permutation-fair (resp. go-first-fair)
    formula. Then \(F_T\) is permutation-fair (resp. go-first-fair)
    for all \(T\subseteq S\).
\end{proposition}

\begin{theorem}
    Suppose \(F:n\to S\cup\{X\}\) is a go-first-fair formula such that 
    \(F_S\) is permutation fair, and for each \(m\leq n\) where \(f(m)=X\), 
    \((F\rest m)_S\) is permutation-fair.
    Then \(F\) is permutation-fair.
\end{theorem}

@@

    So let \(p,q\) be permutations of \(S\cup\{X\}\); we aim to show that
    \(|\{s\in\sigma (F):r_s=p\}|=|\{s\in\sigma (F):r_s=q\}|\).

    First, suppose \(p(|S|)=q(|S|)=X\). Thus 
    \[
        |\{s\in\sigma (F):r_s=p\}|
    =
        \sum\Big\{\Big|\{s\in\sigma ((F\rest m)_S):r_s=p\rest|S|\}\Big|:f(m)=X\Big\}
    \] 
    and likewise for \(q\). Since \((F\rest m)_S\) is permutation-fair,
    \[
        |\{s\in\sigma ((F\rest m)_S):r_s=p\rest|S|\}|
    =
        |\{s\in\sigma ((F\rest m)_S):r_s=q\rest|S|\}|
    \]
    so the result follows in this case.
